\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{A Radial Time Projection Chamber for $\alpha$ detection in CLAS at JLab}

### 2017-04-04 eg6-letters/NIM/EG6_RTPC_NIM.tex
Corrections of section 2

\section{CLAS-EG6 RTPC design} \label{sec_design}
The design of CLAS-EG6 RTPC has been chosen according to physics and 
experimental requirements.  At 6 GeV incident electron beam energy, recoiling 
$^{4}$He nuclei from coherent DVCS have an average momentum (per charge) around
125 MeV/c, while the ``standard'' CLAS detects charged particles above a threshold
of 250 MeV/c. The chosen design also has to fit within the available space inside
a 5 Tesla solenoid magnet surrounding the target \cite{Hyon-suk}. Figure
\ref{fig:RTPC2} shows a photograph of the RTPC before being inserted into the solenoid.

The CLAS-EG6 RTPC is a 200~mm long cylinder with a diameter of 150~mm. As charged 
particles traverse the sensitive gas volume, ionized electrons are released and
drift, under the effect of an applied electromagnetic field, outward perpendicularly
to the beam direction. These electrons are then amplified by three layers of GEMs
and detected by a curved readout system on the external shell, as can be seen in
the figure~\ref{fig:RTPC_1_4}. The EG6-RTPC has two detection halves, but, unlike
the BoNuS RTPC, they share a single chamber.  Each half has independent GEM
amplification systems that cover about 80\% of the azimuthal angle.

\begin{figure}[tb]
\centering
\includegraphics[scale=0.19]{fig/RTPC_exp.png}
\caption{View of the RTPC ready to be inserted in the solenoid. } 
\label{fig:RTPC2}
\end{figure}

\begin{figure}[tb]
\centering
\includegraphics[scale=0.28]{fig/RTPC_1_all.png}
\caption{Schematic drawing of CLAS-EG6 RTPC in a plane perpendicular to 
the beam direction.  See text for description of the elements.} 
\label{fig:RTPC_1_4}
\end{figure}

As seen in figure~\ref{fig:RTPC_1_4}, the CLAS-EG6 RTPC has the following 
substructure, from the beam line to the exterior:\\
\begin{itemize}
   \item The target extends along the RTPC's central z-axis, with a diameter of 
      6 mm and enclosed by a 27-$\mu$m-thick Kapton wall.

   \item The first gas gap covers a radial range of 3 mm to 20 mm. It is 
      filled with $^{4}$He gas at 1~atm to minimize secondary interactions induced
      by M\o ller electrons scattered by the beam. This 
      region is surrounded by a 4 $\mu$m thick window made of aluminized Mylar 
      and electronically grounded.                                                                
   \item The second gap region extends between 20~mm and 30~mm and is filled with a 
      gas mixture of 80$\%$ Neon and 20$\%$ Dimethyl Ether (DME). This region 
      is surrounded by a 4~$\mu$m thick window made of aluminized Mylar, which 
      serves as the cathode.

   \item The drift region is filled with the same Ne-DME gas mixture and extends 
      from the cathode to the first Gas Electron Multiplier(GEM), 60 mm away 
      from the beam axis. The electric field in this region is around 500~V/cm 
      (???) and perpendicular to the beam. Field cages are installed at both 
      ends of the detector in this region. Progressive voltage depending on 
      radial position is applied on these to maintain a uniform electric field 
      perpendicular to the beam line at the edges of the RTPC.

   \item The electron amplification system is composed of three GEMs located at 
      radii of 60, 63 and 66 mm. In this configuration, the first GEM layer 
      represents the anode.

   \item The readout board has an internal radius of 69~mm and collects charge
     from the GEMs. Preamplifiers are plugged directly on its outer side and
     transmit the signal to the data acquisition electronics.
\end{itemize}

\begin{figure}[tbp]
\centering
\includegraphics[scale=0.70]{fig/GEM_photo.jpg}
\caption{A schematic layout of the GEMs system in one module of the RTPC.} 
\label{fig:GEMs}
\end{figure}

The GEM system has been mainly chosen for this detector for its flexibility, 
allowing a curved amplification surface. Also, GEMs are  known to create less 
sparks than other amplicifation technologies for gaseous detectors, which
provides a cleaner detection of slow nuclei like $^4$He. The GEMs for the EG6-RTPC
are made from a Kapton insulator layer sandwiched between two 5 $\mu$m copper
layers. The mesh of each GEM layer is chemically etched with 50$\mu$m diameter
holes with double-conical shapes as can be seen in figure~\ref{fig:GEMs}. A
potential difference of 400 V is applied between the two copper layers of each
GEM, creating a very strong electric field leading to high ionization 
from initial electrons and therefore amplification of the signal. Also, a 
potential difference of 150 V is set between the GEMs to push the amplified 
electrons toward the readout pads. The gain of each GEM layer is of order 100, 
making a total gain of about $10^{6}$.\\

\section{Readout System} \label{sec_readout}
The RTPC electron collection system has 3200 readout pads. These elements are
located at the end of the amplification region, 69 mm from the central axis.

>>>>

\section{CLAS RTPC design} \label{sec_design}

With a 6 GeV incident electron energy, the recoiling $^{4}$He nuclei from coherent 
DVCS have an average momentum around 250 MeV/c. Such low energy $\alpha$ 
particles are stopped very rapidly, so the RTPC was designed to be as close 
as possible to the target and fitting within the 230~mm diameter 
of the 5 Tesla solenoid magnet surrounding the target \cite{Hyon-suk}.

\begin{figure}[tb]
\centering
\includegraphics[scale=0.28]{fig/RTPC_1_all.png}
\caption{Schematic drawing of CLAS RTPC in a plane perpendicular to 
the beam direction.  See text for description of the elements.} 
\label{fig:RTPC_1_4}
\end{figure}

The CLAS RTPC is a 250~mm long cylinder of 158~mm diameter, leaving just enough 
room to fit pre-amplifiers between the RTPC outer shell and the solenoid. The 
electromagnetic field is directed perpendicularly to the beam direction, 
such that drifting electrons are pushed away from the beam line. These electrons 
are amplified by three layers of curved GEMs and detected by the readout system 
on the external shell of the detector as illustrated in the figure~\ref{fig:RTPC_1_4}. The CLAS 
RTPC is segmented into two halves with independent GEM amplification systems that 
cover about 80\% of the azimuthal angle.

We detail here the different regions shown in figure~\ref{fig:RTPC_1_4} starting 
from the beam line and towards larger radius:\\
\begin{itemize}
   \item First, the 6 atm helium-4 target extends along the beam line axis, it 
      is 284~mm long with a 3~mm radius wall made of a 27-$\mu$m-thick Kapton.
   \item The first gas gap covers a radial range from 3 mm to 20 mm. It is 
      filled with $^{4}$He gas at 1~atm to minimize secondary interactions from
      M\o ller electrons scattered by the beam. This 
      region is surrounded by a 4 $\mu$m thick window made of aluminized Mylar 
      and electronically grounded.
   \item The second gap region extends between 20~mm and 30~mm and is filled with a 
      gas mixture of 80$\%$ Neon and 20$\%$ Dimethyl Ether (DME). This region 
      is surrounded by a 4~$\mu$m thick window made of aluminized Mylar, which 
      serves as the cathode. {\bf \color{red} What Voltage ?}
   \item The drift region is filled with the same Ne-DME gas mixture and extends 
      from the cathode to the first Gas Electron Multiplier(GEM), 60 mm away 
      from the beam axis. The electric field in this region is around 500~V/cm 
      {\bf \color{red} Give exact value.} and perpendicular to the beam.
   \item The electron amplification system is composed of three GEMs located at 
      radii of 60, 63 and 66 mm. In this configuration, the first GEM layer 
      represents the anode. {\bf \color{red} What bias voltage ?}
   \item The readout board has an internal radius of 69~mm and collects charge
     from the GEMs. Preamplifiers are plugged directly on its outer side and
     transmit the signal to the data acquisition electronics.
\end{itemize}

\begin{figure}[tbp]
\centering
\includegraphics[scale=0.70]{fig/GEM_photo.jpg}
\caption{Picture of a GEM foil similar to the one used for our RTPC.} 
\label{fig:GEMs}
\end{figure}

The GEM technology has been chosen for this RTPC for its flexibility, 
allowing a curved amplification surface. Also, GEMs are known to rarely create 
sparks{\bf \color{red} Need reference.}, which is important when trying to 
detect highly ionizing slow nuclei that deposit large amount of energy. The 
GEMs for this RTPC are made from a Kapton insulator layer 
{\bf \color{red} Thickness?} sandwiched between two 5 $\mu$m copper
layers. The mesh of each GEM layer is chemically etched with 50$\mu$m diameter
holes with double-conical shapes as illustrated in figure~\ref{fig:GEMs}. A
potential difference of 400 V {\bf \color{red} To be checked and set as above.} 
is applied between the two copper layers of the
GEM, creating a very strong electric field in each hole leading to high ionization 
and therefore amplification of the incoming electrons. A 
potential of 150 V {\bf \color{red} To be checked.} 
is set between the GEMs to push the electrons toward the next GEM and eventually 
the readout pads. The gain of each GEM layer is of order 100, making a total gain 
of about $10^{6}$. {\bf \color{red} We should quote a realistic overall gain or nothing.}\\

{\bf \color{red} We should shortly discuss the choice of drift gas.}

\section{Readout System} \label{sec_readout}

The RTPC electron collection system has 3200 readout pads. These elements are
located at the end of the amplification region, 69 mm from the central axis.
